---
title: "Veg only code for sophie"
author: "Jacob Weverka"
date: "9/6/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(terra)
library(prism)
library(sp) 
library(janitor)
library(pacman)
library(tidyterra)
library(spData)
library(lubridate)

```



# Read in RMN geographic data

```{r include = FALSE}
point_files = list.files(here("data/locations"))

all_points = tibble(point_files) %>% 
  mutate(data = map(point_files, ~ read_csv(file.path(here("data/locations", .x))))) %>% 
  unnest(data) %>% 
  clean_names() %>% 
  filter(types == "Point Count Point")

points_vector = all_points %>% 
  select(name, latitude, longitude) %>% 
  vect(geom = c("longitude", "latitude"))

crs(points_vector) = "EPSG:4326"

plot(points_vector)

```


# Read in Veg Data

## LPI


This reads in the the LPI data. Since you'll have it downloaded in several distinct CSVs from CADC, just put all the lpi CSVs together in a folder. The object `lpi_files` is just the name of the directory folder where these CSVs live. Then `lpi_data_raw` is the result of reading in all of these CSVs and sticking them together.

```{r include=FALSE}

lpi_files = list.files(here("data" , "veg", "lpi"))

lpi_data_raw = tibble(filename = lpi_files) %>% 
  mutate(data = map(filename, ~ read_csv(here("data", "veg", "lpi", .x)))) %>% 
  unnest(data) %>% 
  clean_names() 


```


Some cleaning and summarizing of LPI data: `lpi_count` just counts the number of pin drops at each point. `lpi_sum` calculates the cover of each species in each point - displayed long form. `com_table` contains this same data, but in wide community table format. `species_diversity` contains Shannon-wiener diversity index calculation for each point in each year. `lpi_fun_sum` is like `lpi_sum` except species have been lumped into functional groups.

```{r}

lpi_count = lpi_data_raw %>% 
  filter(!is.na(soil_surface)) %>% 
  group_by(point_id, date) %>% 
  count() %>% 
  mutate(year = lubridate::year(date))


lpi_sum = lpi_data_raw %>% 
  filter(!is.na(soil_surface)) %>% 
  mutate(date = lubridate::ymd(date),
         year = lubridate::year(date)) %>% 
  select(point_id, point_index, year, top_layer, starts_with("lower"), soil_surface) %>% 
  janitor::remove_empty(which = "rows") %>% 
  pivot_longer(c(top_layer, starts_with("lower"), soil_surface)) %>% 
  filter(!is.na(value)) %>% 
  group_by(point_id, year, point_index) %>% 
  select(-name) %>% 
  distinct() %>% 
  ungroup() %>% 
  group_by(point_id, year) %>% 
  count(value) %>% 
  filter(!value %in% c("L", "S", "AM", "R", "M", "NOPLANT", "EM")) %>%
  left_join(lpi_count, by = c("point_id", "year")) %>% 
  group_by(point_id, year) %>% 
  mutate(cover = n.x/sum(n.x))

com_table = lpi_sum %>% 
  filter(!value %in% c("L", "S", "AM", "R", "M", "NOPLANT", "EM")) %>% 
  pivot_wider(id_cols = c(point_id, date, year), names_from = value, values_from = cover)

species_diversity = com_table %>% 
  mutate(name_date = paste(point_id, date, sep = ";")) %>% 
  column_to_rownames("name_date") %>% 
  select(-point_id, -date, -year) %>% 
  mutate_all(~ replace_na(.x, 0)) %>% 
  diversity() %>% 
  as.data.frame() %>% 
  rownames_to_column() %>%
  mutate(point_id = str_sub(rowname, start = 1, end = -12),
         date = ymd(str_sub(rowname, start = -10, end = -1))) %>% 
  clean_names() %>% 
  select(point_id, date, x) %>% 
  mutate(year = year(date))


load(here("data/veg/CAPlantsv2.RData"))

fungrps = CAPlantsv2 %>% 
  janitor::clean_names() %>% 
  select(symbol, fun_grp)

lpi_fun_sum = lpi_sum %>% 
  left_join(fungrps, by = c("value" = "symbol"))



lpi_fun_sum$fun_grp[lpi_fun_sum$value == "2GA"] = "Annual Grass"
lpi_fun_sum$fun_grp[lpi_fun_sum$value == "2FA"] = "Annual Forb"
lpi_fun_sum$fun_grp[lpi_fun_sum$value == "2GP"] = "Perennial Grass"

```


## Unidentified functional groups - a lot of these might end up getting addressed on a case-by-case basis

```{r}
un = lpi_fun_sum %>% 
  filter(is.na(fun_grp),
         !value %in% c("L", "M", "R", "S", "WL", "NOPLANT", "2PLANT", "AM") )
```

`com_table_fun` makes a community table with functional groups


```{r}

com_table_fun = lpi_fun_sum %>% 
  group_by(point_id, year) %>% 
  mutate(rc = n.x/sum(n.x)) %>% 
  group_by(point_id, year, fun_grp) %>% 
  summarise(n = sum(rc)) %>% 
  pivot_wider(id_cols = c(point_id, year), names_from = fun_grp, values_from = n) %>% 
  mutate_all(~ replace_na(.x, 0)) %>% 
  select(-`NA`) %>% 
  clean_names()

```


## Tree demography data

Similar to reading in LPI data


```{r include = FALSE}

dbh_files = list.files(here("data" , "veg", "visit"))

dbh_data_raw = tibble(filename = dbh_files) %>% 
  mutate(data = map(filename, ~ read_csv(here("data", "veg", "visit", .x)))) %>% 
  unnest(data) %>% 
  clean_names() %>% 
  mutate(year = lubridate::year(date))


```



## Releve data

Similar to reading in LPI data

```{r include = FALSE}

releve_files = list.files(here("data", "veg", "releve"))

releve_data_raw = tibble(filename = releve_files) %>% 
  mutate(data = map(filename, ~ read_csv(here("data", "veg", "releve", .x))))


```
